---
title: "Univariate Viz"
---

Use this file for practice with the **univariate viz** in-class activity.  Refer to the class website for details.

# Load the package
library(tidyverse)
# Load the package
library(tidyverse)



library(tidyverse)
# after adding ggplot(hikes, aes(x = rating)) + 
#+       geombar() the count of each difficulty level was added. also the difficulties have been in alphabetical order
# after adding ggplot(hikes, aes(x = rating)) +
#+     geombar() +
#      labs(x = "Rating", y = "Number of hikes") the plot now has labels for its axis

# after adding ggplot(hikes, aes(x = rating)) +
#+     geom_bar(fill = "blue") +
#+     labs(x = "Rating", y = "Number of hikes")
# the bars on the plot are now blue
# after adding ggplot(hikes, aes(x = rating)) +
#+     geom_bar(color = "orange", fill = "blue") +
#+     labs(x = "Rating", y = "Number of hikes")
# the borders of the bars on the plot are now orange

#AFTER ADDING  ggplot(hikes, aes(x = rating)) +
#+     geom_bar(color = "orange", fill = "blue")  +
#+     labs(x = "Rating", y = "Number of hikes") +
#+     theme_minimal()
# the plot now has lines extending from the axis into the plot. I am not good at describing, it is very much in the same spirit of a waffle thematically
#The plot with x as the quantitative variable representing Elevation may not be effective becasue the graph has very extremely thin bars and you cannot see the exact elevation number
#The change in the code is that instead of geom_bar it is now using geom_histogram() 
#The code has now added the color white for the histogram to use which creates the lines that extend from the axis and provides measurements. 
#now the histogram's bars are filled with blue, the structure is pretty much the same aside from the previously mentioned change
#The code has subtracted the part where the bars are filld with the blue color, there are now labels for the axis
#The binwidth is changed to 1000. Wow this is awesome! 
#The bins are now changed to the width of 5! The lines are as thin as hair!
#The bin width is now 200. This is a very pleasant on the eyes and looks like a normal histogram.

#IMPORTANT NOTE: I was using the console to note my work when I should have used this because the console does not save. Below i have included the commands i use.


```{r}

# Load the package
library(tidyverse)

# Import data
hikes <- read.csv("https://mac-stat.github.io/data/high_peaks.csv")

head(hikes)

# Use the ggplot function
ggplot(hikes, aes(x = rating))


# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = rating)) +
  geom_bar()


# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = rating)) +
  geom_bar() +
  labs(x = "Rating", y = "Number of hikes")

# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = rating)) +
  geom_bar(fill = "blue") +
  labs(x = "Rating", y = "Number of hikes")

# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = rating)) +
  geom_bar(color = "orange", fill = "blue") +
  labs(x = "Rating", y = "Number of hikes")

# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = rating)) +
  geom_bar(color = "orange", fill = "blue")  +
  labs(x = "Rating", y = "Number of hikes") +
  theme_minimal()

ggplot(hikes, aes(x = elevation))


# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = elevation)) +
  geom_histogram()

# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white") 

# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", fill = "blue") 



# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white") +
  labs(x = "Elevation (feet)", y = "Number of hikes")

# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", binwidth = 1000) +
  labs(x = "Elevation (feet)", y = "Number of hikes")

# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", binwidth = 5) +
  labs(x = "Elevation (feet)", y = "Number of hikes")

# COMMENT on the change in the code and the corresponding change in the plot
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", binwidth = 200) +
  labs(x = "Elevation (feet)", y = "Number of hikes")




```
```{r}
ggplot(hikes, aes(x = elevation)) +
  geom_density()
```

ggplot(hikes, aes(x = elevation)) +
  geom_density()
  ggplot(hikes, aes(x = elevation)) +
  geom_density()
  
  ggplot(hikes, aes(x = elevation)) +
  geom_density()
  
  
  
```{r}

ggplot(hikes, aes(x = elevation)) +
  geom_density(color= "blue" , fill= "orange" )




```


Good note: make a line break after each + and an indent after the first line of code

recovered ica 1
