---
title: "Bivariate Viz"
---

Use this file for practice with the **bivariate viz** in-class activity.  Refer to the class website for details.
# Import data
survey <- read.csv("https://hash-mac.github.io/stat112site-s25/data/survey.csv")


```{r}

survey <- read.csv("https://hash-mac.github.io/stat112site-s25/data/survey.csv")

```



# How many students have now filled out the survey?
49 students 

# What type of variables do we have?
categorical and quantitative


# Attach a package needed to use the ggplot function

```{r}

library(ggplot2)



```




# Make a ggplot


```{r}


ggplot(survey, aes(x = hangout )) +
  geom_bar(color = "orange", fill = "blue")  +
  labs(x = "Hangout location", y = "Students") +
  theme_minimal()
  
  
```

Now I am making and interpreting two graphs of the students favorite temperature in celcius.
```{r}

ggplot(survey, aes(x = fav_temp_c )) +
  geom_histogram(color = "white", binwidth = 10) +
  labs(x = "Favorite temperature (Celcius)", y = "Number of students")



```

```{r}

ggplot(survey, aes(x = fav_temp_c)) + 
  geom_density(color = "blue", fill = "orange") + 
  labs(x = "favorite temperature C", y = "Students")


```

The plots are unimodal with a peak around 23 celcius. Both have outliers that are over 50 and below around 10 celcius.The data is skewed left with a leftward tail.

