---
title: "Spatial Viz"
---

Use this file for practice with the **spatial viz** in-class activity.  Refer to the class website for details.
```{r}

# Load tidyverse package for plotting and wrangling
library(tidyverse)

# Import the data and only keep 2 addresses
energy <- read.csv("https://mac-stat.github.io/data/MacNaturalGas.csv") |> 
  mutate(date = as.Date(paste0(month, "/1/", year), "%m/%d/%Y")) |> 
  filter(address != "c")

# Check it out
head(energy)

```

```{r}

library(codebook)

```
```{r}

ggplot(energy, aes(x = therms, fill = address)) +
         geom_density(alpha = 0.5)
```


There is a range from a little over 450 to 0 in therms. Address B is more represented in terms of low spending


```{r}

ggplot(energy, aes(x = date, y = therms, color = address)) +
  geom_line() 


```
Aporaching the end of the year the prices rise, at the start of the year the prices decrease until halfway through, where they plateu, and then they increase approaching the end of the year again. This is most likely in line with the rising and decreasing temperatures of the seasons.



Now we will examine to see how rennovation changes therms


```{r}

ggplot(energy, aes(x = renovated, y = therms)) + 
    geom_boxplot() +
    facet_wrap(~ address)
  


```

It looks like rennovation increased the amount of thermal energy used.



```{r}

# A density plot isn't very helpful for comparing typical therms in this example!
ggplot(energy, aes(x = therms, fill = renovated)) + 
  geom_density(alpha = 0.5) + 
  facet_wrap(~ address)


```
I am going to try control for the outside temperature. Maybe the surveyed renovated properties are from colder places.


```{r}

ggplot(energy, aes(y = hdd, x= renovated)) +
  geom_boxplot() + 
  facet_wrap(~ address)


```
Most of the renovated properties are at places with a higher hdd, meaning that they are at colder temperatures and need more therms to be heated.


Now we will make a plot that controls for hdd

```{r}

# When controlling for outside temps (via hdd), energy use decreased post-renovation
ggplot(energy, aes(y = therms, x = hdd, color = renovated)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~ address)
  

```

It appears there is less energy use in the renovated homes

```{r}

install.packages("leaflet")

```
```{r}

library(leaflet)

```
```{r}

fave_places <- read.csv("https://hash-mac.github.io/stat112site-s25/data/our_fave_places.csv")

# Check it out
head(fave_places)

```
```{r}

# Load the leaflet package
library(leaflet)

# Just a plotting frame
leaflet(data = fave_places)

```
```{r}

# Now what do we have?
leaflet(data = fave_places) |> 
  addTiles()

```


```{r}

# Now what do we have?
# longitude and latitude refer to the variables in our data
leaflet(data = fave_places) |> 
  addTiles() |> 
  addMarkers(lng = ~longitude, lat = ~latitude)

```

Awesome we have markers based on latitude and longitude on a map that we can zoom in and out of!!!!! :3


```{r}

# Since we named them "longitude" and "latitude", the function
# automatically recognizes these variables. No need to write them!
leaflet(data = fave_places) |> 
  addTiles() |> 
  addMarkers()


```
```{r}

install.packages("gplots")

```

```{r}

library(gplots)


```
```{r}

# Load package needed to change color
library(gplots)

# We can add colored circles instead of markers at each location
leaflet(data = fave_places) |> 
  addTiles() |> 
  addCircles(color = col2hex("red"))


```



```{r}

# We can change the background
# Mark locations with yellow dots
# And connect the dots, in their order in the dataset, with green lines
# (These green lines don't mean anything here, but would if this were somebody's travel path!)
leaflet(data = fave_places) |>
  addProviderTiles("USGS") |>
  addCircles(weight = 10, opacity = 1, color = col2hex("yellow")) |>
  addPolylines(
    lng = ~longitude,
    lat = ~latitude,
    color = col2hex("green")
  )


```


```{r}

# Import starbucks location data
starbucks <- read.csv("https://mac-stat.github.io/data/starbucks.csv")

```

```{r}

# Don't worry about the syntax
starbucks_mn <- starbucks |>   
  filter(Country == "US", State.Province == "MN")


```

```{r}

leaflet(starbucks_mn) |> 
  addMarkers() |> 
  addTiles() 


```

```{r}


ggplot(starbucks, aes(x = Longitude, y = Latitude)) +
  geom_point(size = 0.2, alpha = 0.2)

```


```{r}
install.packages("rnaturalearth")
```

```{r}

# Load the package
library(rnaturalearth)

# Get info about country boundaries across the world
# in a "sf" or simple feature format
world_boundaries <- ne_countries(returnclass = "sf")

```

```{r}

# What does this code produce?
# What geom are we using for the point map?
ggplot(world_boundaries) + 
  geom_sf()

```

```{r}

# Load package needed to change map theme
library(mosaic)

# Add a point for each Starbucks
# NOTE: The Starbucks info is in our starbucks data, not world_boundaries
# How does this change how we use geom_point?!
ggplot(world_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3, size = 0.2, color = "darkgreen"
  ) +
  theme_map()



```

According to this map, starbucks has a multinational presence, especially in the US on the coasts, in Japan, in the UK, in China, in France and also in Germany.


```{r}

# We'll learn this syntax soon! Don't worry about it now.
starbucks_cma <- starbucks |> 
  filter(Country %in% c('CA', 'MX', 'US'))




```



```{r}

install.packages("rnaturalearthdata")

```

```{r}

install_rnaturalearthhires()


```


```{r}


```

```{r}

install.packages("devtools") # I guess you also need this
devtools::install_github("ropensci/rnaturalearthhires")
library("rnaturalearth")

```
I looked up how to fix a bug that would not let me use the next code and it looks like the remedy was installing devtools in order to manually run the code, hope this works!!

```{r}

remotes::install_github("ropensci/rnaturalearthhires")




```
```{r}

     devtools::install_github("ropensci/rnaturalearthhires")

```
```{r}

force = TRUE

```

```{r}

library(rtools)

```

```{r}

install_rnaturalearthdata()

```
```{r}

 install.packages("rnaturalearthdata")

```
```{r}

library(rnaturalearthhires)


```


```{r}

cma_boundaries <- ne_states(
  country = c("canada", "mexico", "united states of america"),
  returnclass = "sf")

```

```{r}

# We'll learn this syntax soon! Don't worry about it now.
starbucks_cma <- starbucks |> 
  filter(Country %in% c('CA', 'MX', 'US'))

```


```{r}


# Just the boundaries
ggplot(cma_boundaries) + 
  geom_sf()

```
```{r}

# Add the points
# And zoom in
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3,
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50)) +
  theme_map()


```




```{r}



starbucks_midwest <- starbucks |> 
  filter(State.Province %in% c("MN", "ND", "SD", "WI"))

```


```{r}

install.packages("maps")

```


```{r}

# Load packages
library(sf)
library(maps)

# Get the boundaries
midwest_boundaries <- st_as_sf(
  maps::map("county",
            region = c("minnesota", "wisconsin", "north dakota", "south dakota"), 
            fill = TRUE, plot = FALSE))

# Check it out
head(midwest_boundaries)

```

```{r}

# Load packages
library(sf)
library(maps)

# Get the boundaries
midwest_boundaries <- st_as_sf(
  maps::map("county",
            region = c("minnesota", "wisconsin", "north dakota", "south dakota"), 
            fill = TRUE, plot = FALSE))

# Check it out
head(midwest_boundaries)


```
```{r}


 ggplot(midwest_boundaries) + 
   geom_sf() + 
   geom_point(
     data = starbucks_midwest,
     aes(x = Longitude, y = Latitude),
     alpha = 0.7,
     size = 0.2, 
     color = 'darkgreen'
   ) + 
   theme_map()

```
```{r}

# Point map (we made this earlier)
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3,
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50), ylim = c(14.54, 83.11)) +
  theme_map()


```

```{r}

# What changed in the plot?
# What changed in our code?!
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_density_2d(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50), ylim = c(14.54, 83.11)) +
  theme_map()



```

```{r}

elections_by_state <-  read.csv("https://mac-stat.github.io/data/election_2020_by_state.csv")
elections_by_counties <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")


```

```{r}


# Don't worry about the code!

elections_by_state <- elections_by_state |> 
  filter(state_abbr != "DC") |> 
  select(state_name, state_abbr, repub_pct_20) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(30, 70, by = 5), 
               labels = c("30-34", "35-39", "40-44", "45-49",
                          "50-54", "55-59", "60-64", "65-70"), 
               include.lowest = TRUE))

elections_by_counties <- elections_by_counties |> 
  select(state_name, state_abbr, county_name, county_fips,
          repub_pct_20, median_age, median_rent) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(0, 100, by = 10),
               labels = c("0-9", "10-19", "20-29", "30-39", "40-49",
                          "50-59", "60-69", "70-79", "80-89", "90-100"),
               include.lowest = TRUE))

```


```{r}

# Get the latitude and longitude coordinates of state boundaries
states_map <- map_data("state")

# Check it out
head(states_map)


```

```{r}

head(states_map)

```

```{r}

head(elections_by_state) 

```

```{r}


# Note where the dataset, elections_by_state, is used
# Note where the background map, states_map, is used
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() 

```

```{r}


# Make it nicer!
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_gradientn(name = "% Republican", colors = c("blue", "purple", "red"), values = scales::rescale(seq(0, 100, by = 5)))

```

```{r}

ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map()


```
```{r}

install.packages("RColorBrewer")

```
```{r}

# Load package needed for refining color palette
library(RColorBrewer)

# Now fix the colors
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_manual(values = rev(brewer.pal(8, "RdBu")), name = "% Republican")

```

```{r}


# Get only the starbucks data from the US
starbucks_us <- starbucks |> 
  filter(Country == "US")

# Map it
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  geom_point(
    data = starbucks_us,
    aes(x = Longitude, y = Latitude),
    size = 0.05,
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_manual(values = rev(brewer.pal(8, "RdBu")), name = "% Republican")

```

Looks like starbucks correlate with population. Still pretty cool though!!!!!


```{r}


install.packages("socviz")


```
```{r}

# Get the latitude and longitude coordinates of county boundaries
library(socviz)
data(county_map) 

# Check it out
head(county_map)

```
```{r}
head(county_map)


```
```{r}

head(elections_by_counties)

```
```{r}

# Add 0's at the beginning of any fips_code that's fewer than 5 numbers long
# Don't worry about the syntax
elections_by_counties <- elections_by_counties |> 
  mutate(county_fips = as.character(county_fips)) |> 
  mutate(county_fips = 
           ifelse(nchar(county_fips) == 4, paste0("0", county_fips), county_fips))


```

```{r}

ggplot(elections_by_counties, aes(map_id = county_fips, fill = repub_20_categories)) +
  geom_map(map = county_map) +
  scale_fill_manual(values = rev(brewer.pal(10, "RdBu")), name = "% Republican") +
  expand_limits(x = county_map$long, y = county_map$lat) +
  theme_map() +
  theme(legend.position = "right") + 
  coord_equal()


```

```{r}
ggplot(elections_by_counties, aes(map_id = county_fips, fill = median_rent)) +
  geom_map(map = county_map) +
  expand_limits(x = county_map$long, y = county_map$lat) +
  theme_map() +
  theme(legend.position = "right") + 
  coord_equal() + 
  scale_fill_gradientn(name = "median rent", colors = c("white", "lightgreen", "darkgreen"))


```

```{r}

ggplot(elections_by_counties, aes(map_id = county_fips, fill = median_age)) +
  geom_map(map = county_map) +
  expand_limits(x = county_map$long, y = county_map$lat) +
  theme_map() +
  theme(legend.position = "right") + 
  coord_equal() + 
  scale_fill_gradientn(name = "median age", colors = c("orange", "lightgreen", "blue"))




```
Spatial ICA complete
